---
title: "Intro to readxl"
author: "Michelle"
date: December 7, 2019
output: html_document
---

# Learning `readxl`

We are working with data in readxl and it's going to be amazing

We are going to use packages:

- tidyverse
- readxl
- writeexl
- here


```{r package}
library(tidyverse)
library(readxl)
library(writexl)
library(here)

# install.packages ("package_name")

```
### Read in fish counts

```{r}
##### Read in the fish_counts_curated.csv
fish_counts <- read.csv(here("R-workshop-data", "fish_counts_curated.csv"))
fish_counts
```

### Read in data from a single Excel worksheet
Now, to read in an excel spreasheet. we'll use the function `read_excel()`
```{r}
substrate_cover <-read_excel(here("R-workshop-data", "substrate_cover_curated.xlsx"),
                             skip=4,
                             na="-9999")
```
### Use read_excel () to read in a range of values

```{r}
carp_cover_2000 <- read_excel(here("R-workshop-data", "substrate_cover_curated.xlsx"), range = "D12:I15", 
                              col_names = c ("site",
                                             "transect", 
                                             "quad",
                                             "side",
                                             "type",
                                             "coverage"))
                                   
```

#### How can we read in data from an Excel workbook with Multiple workbooks?

We only want to read in data for Goleta Beach. Worksheet named golb

```{r}
kelp_golb <- read_excel(here("R-workshop-data","kelp_counts_curated.xlsx"), 
                        sheet = "golb")

```

#### Read in AND combine data from multiple Excel worksheets simultaneously using purrr:::map_df()

```{r}
kelp_path <- here("R-workshop-data", "kelp_counts_curated.xlsx")


```

### Write a final back to an Excel spreadsheet

We're going to write our "kelb_golb" object as a CSV file

```{r}
write_csv(kelp_golb,"kelp_goleta_july.csv")
```


