---
title: "Filter/joins (like VLOOKUP in Excel)"
output: html_document
---

# Attach Packages
tidyverse, readxl, here

```{r, include=FALSE}
library(tidyverse)
library(readxl)
library(here)
```

Read in 3 files:
-invert_counts_curated.xlsx
-fish_counts_curated.csv
-kelp_counts_curated.xlsx

```{r}
invert_counts <- read_excel(here("R-workshop-data", 
                                 "invert_counts_curated.xlsx"))

fish_counts <- read_csv(here("R-workshop-data", "fish_counts_curated.csv"))

  
Kelp_count_abur <- read_excel(here("R-workshop-data", "kelp_counts_curated.xlsx"))
```
### Filter to match rows based on a character string
Only keep observations from fish_counts where the common_name variable matches "garibaldi"
```{r}
fish_garibaldi <- fish_counts %>%
  filter(common_name == "garibaldi")
```

#Use janitor::clean_names () to automatically update all of your column names to lowercase_snake_case

## Filter to keep/exclude observations based on numeric 

```{r}
fish_over50 <- filter(fish_counts, total_count >= 50)
```

```{r}
fish_Mohawk <- fish_counts %>%
  filter(site=="mohk")
```
```{r}
fish_3sp <- fish_counts %>% 
  filter(common_name %in% c("garibaldi", "blacksmith", "black surfperch"))
```

## Filter to return rows that match conditions fo

```{r}
urchin_abur_mohk <- invert_counts %>% 
  filter(site %in% c("abur","mohk"), common_name == "purple urchin")
```

```{r}
low_gb_rw <- fish_counts %>%
  filter(common_name %in% c("garibaldi", "rock wrasse"), 
         total_count<10)
```
Use!= (indicates "does NOT match)

```{r}
fish_no_mohk <- fish_counts %>% 
  filter(site != "mohk")
```

An example of filter as part of a piped sequence

```{r}
senorita_allyears <- fish_counts %>%
  filter(common_name=="senorita") %>% 
  group_by(site) %>% 
  summarize(count_all_years = sum(total_count))
```





### Joining data frames

-'full join' : keeps everything (safe join)
-'left join' : If we are merging A and B, will only retain observations from B that have matches in A


```{r}
fish_2016_2017 <- fish_counts %>%
  filter (year ==2016 | year == 2017)
```


