---
title: "intro-to-ggplot2"
output: html_document
---



## Attach packages

```{r}
library(tidyverse)
library(readxl)
library(here)
```

```{r}
cap_np <- read_csv(here("R-workshop-data", "ca_np.csv"))
ci_np <-read_xlsx(here ("R-workshop-data", "ci_np.xlsx"))
```

## Make a graph

1. Tell R we are using 'ggplot'
2. Tell R what data we are using and what variables we want to plot
3. What type of graph we are creating

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()
```
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point()
```
1. Create a column graph with this data
2. Create an area graph with this data

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_col()
```
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_area()
```
### Intro to customizing 'ggplot' graphs

- color=(update point or line colors)
-fill = (update the fill characteristic of an element)
- linetype
-pch = (value)
-size
-alpha

```{r}
ggplot(data=ci_np, aes(x=year, y=visitors)) + geom_line(color="coral", linetype="dashed")
```
```{r}
ggplot(data=ci_np, aes(x=year, y=visitors)) +
  geom_point(color = "blue",
             size = 4,
             alpha = 0.5)
```

```{r}
ggplot(data=ci_np, aes(x=year, y=visitors)) +
  geom_point(color = "coral",
             size = 3,
             alpha = 1)
```
I have text here and a graph below

```{r, echo=FALSE}
ggplot(data=ci_np, aes(x=year, y=visitors)) +
  geom_point(color = "coral",
             size = 3,
             alpha = 1)
```
## Customize elements based on variable values or outcomes

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_point(aes (size = visitors, 
color = visitors                                                                    ), pch = 17)
```
```{r}
```


```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_line(size = 3, color="red") + 
  theme_minimal()
```


```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_line(size = 3, color="red") + 
  theme_minimal()
```



### Combine compatible geoms

Example : Add a point plot on top of a line plot:

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_line(color="purple") + geom_point(color= "magenta",
                                         aes(size=year))
```
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_col(color="purple") + geom_line(color= "magenta",
                                         aes(size=year))
```
### Create a line graph with multipe series
```{r}
ggplot(data = cap_np, aes(x = year, y = visitors, color = park_name)) + geom_line()
```
### Faceting GG plot graphs
-Each series (group) gets its own plot panel

```{r}
ggplot(data = cap_np, aes(x = year, y = visitors, color = park_name)) + geom_line(show.legend = FALSE) +
  theme_minimal() +
facet_wrap (~ park_name)
```
## Use ggsave to export a plot

```{r}
ggsave("np_graph.png", dpi = 180, width = 8, height = 7)
```

