---
title: "intro-to-ggplot2"
output: html_document
---



## Attach packages

```{r}
library(tidyverse)
library(readxl)
library(here)
```

```{r}
cap_np <- read_csv(here("R-workshop-data", "ca_np.csv"))
ci_np <-read_xlsx(here ("R-workshop-data", "ci_np.xlsx"))
```

## Make a graph

1. Tell R we are using 'ggplot'
2. Tell R what data we are using and what variables we want to plot
3. What type of graph we are creating

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()
```
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point()
```
1. Create a column graph with this data
2. Create an area graph with this data

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_col()
```
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_area()
```
### Intro to customizing 'ggplot' graphs

- color=(update point or line colors)
-fill = (update the fill characteristic of an element)
- linetype
-pch = (value)
-size
-alpha

```{r}
ggplot(data=ci_np, aes(x=year, y=visitors)) + geom_line(color="coral", linetype="dashed")
```
```{r}
ggplot(data=ci_np, aes(x=year, y=visitors)) +
  geom_point(color = "blue",
             size = 4,
             alpha = 0.5)
```

```{r}
ggplot(data=ci_np, aes(x=year, y=visitors)) +
  geom_point(color = "coral",
             size = 3,
             alpha = 1)
```
I have text here and a graph below

```{r, echo=FALSE}
ggplot(data=ci_np, aes(x=year, y=visitors)) +
  geom_point(color = "coral",
             size = 3,
             alpha = 1)
```

