---
title: "Filter/joins (like VLOOKUP in Excel)"
output: html_document
---

# Attach Packages
tidyverse, readxl, here

```{r, include=FALSE}
library(tidyverse)
library(readxl)
library(here)
```

Read in 3 files:
-invert_counts_curated.xlsx
-fish_counts_curated.csv
-kelp_counts_curated.xlsx

```{r}
invert_counts <- read_excel(here("R-workshop-data", 
                                 "invert_counts_curated.xlsx"))

fish_counts <- read_csv(here("R-workshop-data", "fish_counts_curated.csv"))

  
Kelp_count_abur <- read_excel(here("R-workshop-data", "kelp_counts_curated.xlsx"))
```
### Filter to match rows based on a character string
Only keep observations from fish_counts where the common_name variable matches "garibaldi"
```{r}
fish_garibaldi <- fish_counts %>%
  filter(common_name == "garibaldi")
```

#Use janitor::clean_names () to automatically update all of your column names to lowercase_snake_case

## Filter to keep/exclude observations based on numeric 

```{r}
fish_over50 <- filter(fish_counts, total_count >= 50)
```

```{r}
fish_Mohawk <- fish_counts %>%
  filter(site=="mohk")
```
```{r}
fish_3sp <- fish_counts %>% 
  filter(common_name %in% c("garibaldi", "blacksmith", "black surfperch"))
```

## Filter to return rows that match conditions for MULTIPLE variables (AND statements)

```{r}
urchin_abur_mohk <- invert_counts %>% 
  filter(site %in% c("abur","mohk"), common_name == "purple urchin")
```

