---
title: "Pivot tables"
output: html_document
---
Attach libraries

```{r}
library(tidyverse)
library(readxl)
library(here)
library(skimr) # install.packages ("skimr")
```
Our data is about lobsters
```{r}
lobsters <- read_xlsx(here("R-workshop-data", "lobsters.xlsx"))
```
#head function show the first 6 lines of data
```{r}
head(lobsters)
```
#summary to give a summary of each column
```{r}
summary(lobsters)
```
```{r}
skim(lobsters)
```
```{r}
ggplot(data = lobsters, aes(x=year)) +
  geom_bar() +
  facet_wrap(~site)
```
## Our task is how the average size of lobsters has changed over time across sites
Pivot tables in R are group_by %>% summarize

```{r}
lobsters %>%
  group_by(year) %>%
  summarize(count_by_year = n())
```

```{r}
lobsters %>%
  summarize(count_by_year = n())
## assign our summary to a new variable
lobsters_sum <- lobsters %>%
  summarize(count_by_something = n())
```
```{r}
lobsters %>%
group_by (site, year) %>%
summarize(count_by_siteyear = n())
```
## group_by multilpe variables
```{r}
lobsters %>%
group_by (site, year) %>%
summarize(count_by_siteyear = n(), 
          mean_size_mm=mean(size_mm, na.rm = TRUE))
```
```{r}
siteyear_summary <- lobsters %>%
group_by (site, year) %>%
summarize(count_by_siteyear = n(), 
          mean_size_mm=mean(size_mm, na.rm = TRUE), 
          sd_size_mm = sd(size_mm, na.rm = TRUE))
siteyear_summary
```
```
```{r}
lobsters %>%
group_by (site, year) %>%
summarize(count_by_siteyear = n(), 
          median_size_mm=mean(size_mm, na.rm = TRUE), 
          sd_size_mm = sd(size_mm, na.rm = TRUE))
```

```{r}
lobsters %>%
group_by (site, year) %>%
summarize(count_by_siteyear = n(), 
          median_size_mm=median(size_mm, na.rm = TRUE), 
          sd_size_mm = sd(size_mm, na.rm = TRUE))
```
```{r}
lobsters %>%
group_by (site, year) %>%
summarize(count_by_siteyear = n(), 
          mean_size_mm=median(size_mm, na.rm = TRUE))
```
```{r}
siteyear_summary <- lobsters %>%
group_by (site, year) %>%
summarize(count_by_siteyear = n(), 
          mean_size_mm=mean(size_mm, na.rm = TRUE), 
          sd_size_mm = sd(size_mm, na.rm = TRUE),
          median_size_mm = median (size_mm, na.rm = TRUE))
siteyear_summary

ggplot(data = siteyear_summary, aes (x=year, median_size_mm, color = site)) +
geom_line()
ggsave(here("figures/lobsters-line.png")
```

```{r}
lobsters %>%
group_by (site, year) %>%
summarize(count_by_siteyear = n(), 
          median_size_mm=mean(size_mm, na.rm = TRUE), 
          sd_size_mm = sd(size_mm, na.rm = TRUE))
```
 # dplyr:count() for counting by group
 This is a shortcut function
```{r}
lobsters %>%
count(site,year) %>%
summarize (n=n()) %>%
ungroup()
```

```{r}
head(lobsters)

lobsters %>%
mutate(size_m = size_mm/1000,
millenia = 2000,
observer= "you")
```

let's select 
-date
-site
-size_m

```{r}

lobsters_selected <- lobsters %>%
mutate(size_m = size_mm / 1000) %>%
select(date, site, size_m)
```

 
